Simulating cosmic variance in Weak Lensing: effects on parameter inferences

### 2015-12-03 Paper/paper.tex
finished implementing modifications with O(1/N^2) predictions

\begin{abstract}
Constraining cosmology using Weak Gravitational Lensing consists in comparing measured features with simulated ones. An accurate estimate of the feature covariance matrix is essential to obtain accurate parameter confidence intervals. When this covariance matrix $\bb{C}$ is measured from simulations, an important question to ask is how big the simulation set should be for an accurate estimation of $\bb{C}$. We construct different mock ensembles with $N_r$ realizations of the same shear field, based on $N_s<N_r$ independent $N$--body simulations. Using the shear--shear power spectrum as a summary statistic, we find that a single $N$--body simulation can be recycled a large number of times to produce an ensemble of $N_r=10^5$ shear realizations that are mutually independent.       
\end{abstract}

@@

\subsection{Shear field simulations}
\label{shearsim}
%
In this paragraph we describe how we constructed our shear field ensembles. Background galaxies at redshift $z_s$ are lensed by large scale structure between $z=0$ and $z_s$. The shape distortions due to the cosmic shear $\pmb{\gamma}$ can be computed in terms of the dark matter gravitational potential $\Phi(\bb{x},z)$. Because the evolution of $\Phi$ in redshift is non--linear, its evolution needs to be computed with numerical simulations. We make use of the public code \texttt{Gadget2} \citep{Gadget2}, with which we run a sequence of dark--matter--only $N$--body simulations that track the evolution of the density fluctuations in the large scale structure of the universe. We assume a standard $\Lambda$CDM framework with $(\Omega_m,\Omega_\Lambda,h,w,\sigma_8,n_s)=(0.26,0.74,0.72,-1,0.8,0.96)$ and we fix the comoving size of the simulation box to $240\mathrm{Mpc}/h$. We fill the box with $512^3$ particles, which correspond to a mass resolution of $10^{10}M_\odot$. We assume a uniform galaxy distribution at a constant redshift $z_s=2$ (at which the simulation box has an angular size of $\theta_{box}=3.5\mathrm{deg}$) and we disctretize the mass distribution between $z_s$ and the observer at $z=0$ with a sequence of 46 two dimensional lenses of thickness $80\mathrm{Mpc}/h$. The surface density on each lens is computed by projecting the three dimensional density measured from \texttt{Gadget2} snapshots. We then apply the multi--lens--plane algorithm (see \citep{RayTracingHartlap,RayTracingJain} for example) to trace the deflections of $n_{ray}^2=2048^2$ light rays arranged on a square grid of total size $\theta_{box}$, from $z=0$ to $z_s$. The light deflection calculations have been performed with our implementation of multi--lens--plane algorithm, which is part of the \texttt{LensTools} computing package \citep{LensTools} (released under the \texttt{MIT} license). Different realizations $r$ of the same shear field $\pmb{\gamma}_r(\pmb{\theta})$ can be generated with different choices of the lenses that lay between the observer and $z_s$. The randomization procedure we adopt is the following:

\begin{itemize}
\item Pick a lens redshift $z_l$, and select the snapshot at $z_l$ from the $i$--th $N$--body simulation at disposal, where $i$ is a random number in $[1,N_s]$
\item Choose randomly a direction between the three choices ${\bb{n}_x,\bb{n}_y,\bb{n}_z}$: the lens plane will be pependicular to this direction
\item Choose the cutting point of the plane in the snapshot: because the lens thickness is 1/3 the size of the box, we can cut three different slices of the simulation box for each orientation ${\bb{n}_x,\bb{n}_y,\bb{n}_z}$
\item Perform a periodical random shift of the lens plane along its two directions
\item Repeat this procedure for each lens redshift $z_l$  
\end{itemize}  
%
This randomization procedure allows us to produce an arbitrary number $N_r$ of mock shear realizations $\pmb{\gamma}_r(\pmb{\theta})$, which however will not be all independent if $N_s$ is not large enough. Using a set of 200 independent $N$--body simulations, we construct different ensembles with different choices of $N_s\in[1,200]$, each made of 1000 mock shear realizations. We also build an additional ensemble with $N_s=1$ and $10^5$ mocks. From each of these mocks we reconstruct the convergence $\kappa_r(\pmb{\theta})$ and measure its angular power spectrum $P^{\kappa\kappa}_r(l)$ defined as
\begin{equation}
\langle\tilde{\kappa}_r(\bb{l})\tilde{\kappa}_r(\bb{l}')\rangle = (2\pi)^2\delta_D(\bb{l}+\bb{l}')P^{\kappa\kappa}_r(l)
\end{equation}
%
As an additional summary statistic, we consider the counts of local $\kappa$ maxima of a certain height $\kappa_0$, $n_r(\kappa_0)$ (from now on \textit{peak counts}), with varying $\kappa_0$. The fact that the ensemble of $N_r$ mocks is not completely independent if $N_s$ is not large enough can have an effect on the covariance estimators of $P^{\kappa\kappa},n(\kappa_0)$.

@@

\begin{equation}
\label{covest}
\bbh{C} = \frac{1}{N_r-1}\sum_{r=1}^{N_r} (\bbh{d}_r - \bar{\bb{d}}) (\bbh{d}_r - \bar{\bb{d}})^T
\end{equation}
%
Assuming a normal feature likelihood, together with a flat prior on the parameter space, the parameter posterior distribution $\mathcal{L}(\bb{p}\vert\bbh{d}_{obs})$ given an observed instance of $\bbh{d}$, which we call $\bbh{d}_{obs}$, can be obtained using Bayes theorem
\begin{equation}
\label{posteriorbayes}
-2\log\mathcal{L}(\bb{p}\vert\bbh{d}_{obs}) = (\bbh{d}_{obs}-\bb{d}(\bb{p}))^T\bbh{C}^{-1}(\bbh{d}_{obs}-\bb{d}(\bb{p}))
\end{equation}
%
For the sake of simplicity, we approximate the posterior as a Gaussian around its maximum. This is easily done approximating the simulated feature at first order around a point $\bb{p}_0$ (that ideally is the maximum of (\ref{posteriorbayes}))
\begin{equation}
\bb{d}(\bb{p}) \approx \bb{d}_0 + \bb{d}_0^\prime(\bb{p}-\bb{p}_0) 
\end{equation}
%
To measure the feature derivatives $\bb{d}'_0$ with respect to cosmology, we make use of the public code \texttt{NICAEA} \citep{Nicaea} for the power spectrum, and we use an independent simulation set (that contains mocks at different combinations of $(\Omega_m,w,\sigma_8)$) for the peak counts.
We can build the estimator for the posterior maximum $\bbh{p}$, given the observation $\bbh{d}_{obs}$, as follows:
%
\begin{equation}
\label{estimatormean}
\bbh{p} = \bb{p}_0 + \bbh{T}(\bbh{d}_{obs}-\bb{d}_0)
\end{equation}

@@

\begin{equation}
\label{estimatorfisher}
\bbh{F} = \bb{d}_0'^T\bbh{\Psi}\bb{d}_0'
\end{equation}
%
Where we introduced the Fisher matrix estimator $\bbh{F} = \bb{d}_0'^T\bbh{\Psi}\bb{d}_0'$ and, for simplicity, we assumed $\langle\bbh{d}_{obs}\rangle_O=\bb{d}_0$, so that $\langle(\bbh{d}_{obs}-\bb{d}_0)(\bbh{d}_{obs}-\bb{d}_0)^T\rangle_O=\bb{C}$. 
When we perform an observation $\bbh{d}_{obs}$, the parameter estimate $\bbh{p}$ is a random draw from a probability distribution with variance $\h{\Sigma}_\bb{p}$, which inherits noise from the simulations. In the following we review the method to compute $\langle\h{\Sigma}_\bb{p}\rangle$ \citep{DodelsonSchneider13,Taylor12}. The noise in the covariance estimator (\ref{covest}) and its inverse $\bbh{\Psi}$ propagates all the way to the posterior (\ref{posteriorbayes}), the parameter estimate (\ref{estimatormean}) and its variance (\ref{estimatorcovariance}). Following \citep{Taylor12} we can say that, if the feature vector $\bbh{d}_r$ from which $\bbh{C}$ is estimated is drawn from a multivariate Gaussian distribution, the precision matrix estimator $\bbh{\Psi}$ is biased
\begin{equation}
\label{psibias}
\langle\bbh{\Psi}\rangle = \frac{N_r-1}{N_r-N_b-2}\bb{\Psi}
\end{equation}
%
and, once the bias is corrected, the residual fluctuations in $\bbh{\Psi}$ are correlated as follows
\begin{equation}
\label{psifluctuations}
\langle\delta\h{\Psi}_{ij}\delta\h{\Psi}_{kl}\rangle = A\Psi_{ij}\Psi_{kl} + B(\Psi_{ik}\Psi_{jl} + \Psi_{il}\Psi_{jk})
\end{equation}
%
where $A,B$ are constants that depend on $N_r,N_b$ and have the asymptotic behavior $A\propto1/N_r^2,B\propto1/N_r$. We want to estimate the effect of these fluctuations on the parameter constraints $\Sigma_\bb{p}$, following the guidelines of \citep{DodelsonSchneider13}. We consider three cases:
\begin{enumerate}
\item If the true data covariance $\bb{C}$ is known, the estimator (\ref{estimatorcovariance}) is biased, and the dominant contribution of the bias comes from the second order fluctuations in $\bbh{\Psi}$. Once the expectation values are taken as in (\ref{psifluctuations}), the bias sums up to

\begin{equation}
\label{dodelsonscaling}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}(1+B(N_b-N_p))
\end{equation}
%
This is the result obtained by \citep{DodelsonSchneider13}.

\item Usually the true covariance is not known, and it is tempting to plug in its estimator $\bbh{C}$, measured from the same simulation set we use to compute $\bbh{\Psi}$. This approach has been used before in the literature (see \citep{MinkPetri,MinkShirasaki} for example). If we do that, and we do not correct for the bias (\ref{psibias}), the parameter variance will have a contribution from both the second and first order fluctuations in $\bbh{\Psi}$, which now have a non zero expectation value. In this case the bias sums up to 
\begin{equation}
\label{mockscalinguncorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}(1-(1+N_b)/N_r+B(1+N_p))
\end{equation}

@@

\begin{equation}
\label{mockscalingcorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}(1+B(1+N_p))
\end{equation}

\end{enumerate} 
%
The complete calculations that justify (\ref{dodelsonscaling}),(\ref{mockscalinguncorrected}) and (\ref{mockscalingcorrected}) are shown in the Appendix. With the $\bbh{d}_r$ Gaussianity assumption, the coefficient $B$ can be calculated exactly as \citep{DodelsonSchneider13,Taylor12}, and tends to $1/N_r$ in the limit of $N_r\gg N_b$. This leads to a degradation of order $D/N_r$ for the variance of each estimated parameter $\sigma^2_p=\Sigma_{pp}$, where $D=N_b-N_p,N_p-N_b,1+N_p$ for cases 1,2,3 respectively. This means that, when parameter errors are estimated purely from simulations, and the same simulation set is used to estimate both $\bbh{C},\bbh{\Psi}$, these errors tend to decrease with $N_b$ (case 2), or at best remain constant (case 3), while the distribution of $\bbh{p}$ has still a variance that increases with $N_b$. This typically leads to underestimates in parameter errors. We test the scaling relation
\begin{equation}
\label{ourscaling}
\langle\h{\sigma}_p^2\rangle = \sigma^2_p(N_s)\left(1+\frac{D}{N_r}\right)
\end{equation}
%
on our simulations. We call $D$ the \textit{effective dimensionality} of the feature space. We compute the expectation values of $\h{\sigma}^2_p$ (equations (\ref{estimatorcovariance}),(\ref{ourscaling})) averaging over 100 random resamplings of our shear ensembles. For the true feature covariance matrix $\langle(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\rangle=\bb{C}$ we use the estimated covariance from a grand ensemble built with the union of all the ensembles with different $N_s$.

The true parameter variance $\sigma^2_p(N_s)$ in principle depends on the number of independent $N$--body simulations $N_s$, which appears in the randomization procedure we describe in \S~\ref{shearsim}. The reason for this can be found in that, if $N_s$ is not big enough, the different shear realizations cannot be all independent, and hence the true variance $\sigma^2_p(N_s\rightarrow\infty)$ cannot be recovered in the limit $N_r\rightarrow\infty$. In particular there could be some residual error in the expectation value of (\ref{estimatorcovariance}) even if $N_r$ is big. In the next paragraph we outline our main findings.

@@

@@

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_ns.eps}
\caption{Variance on $w$ in the limit $N_r\rightarrow\infty$, varying the mock ensemble used to compute the covariance estimator (\ref{covest}). We show the dependence of $\sigma_0^2(N_s)$ in units of the mean over $N_s$, for the power spectrum logarithmically binned (black), the power spectrum linearly binned (red) and the peak counts (green).}
\label{wvar_ns}
\end{figure}

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}
\toprule
\textbf{Feature} &  \textbf{Specifications} & $N_b$ &  \textbf{Symbol} & \textbf{Color} \\ \hline \hline
\midrule
Power Spectrum, log binning  & $l \in [100,800] $ & 8 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [1000,6000] $ & 7 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [100,6000] $ & 15 & $\times$ & black  \\
Power Spectrum, linear binning  & $l \in [100,2000] $ & 15 & \textcolor{red}{$+$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $l \in [2500,4500] $ & 15 & \textcolor{red}{$\times$} & \textcolor{red}{red}  \\
Power Spectrum, linear binning  & $l \in [100,4500] $ & 30 & \textcolor{red}{$\bullet$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $l \in [100,6000] $ & 39 & \textcolor{red}{$\bullet$} & \textcolor{red}{red}  \\ \hline
Low peaks  & $\kappa_0 \in [-0.06,0.09] $ & 15 & \textcolor{OliveGreen}{$+$} & \textcolor{OliveGreen}{green}  \\ 
Intermediate peaks  & $\kappa_0 \in [0.1,0.27] $ & 15 & \textcolor{OliveGreen}{$\bigstar$} & \textcolor{OliveGreen}{green}  \\ 
High peaks  & $\kappa_0 \in [0.28,0.45] $ & 15 & \textcolor{OliveGreen}{$\diamond$} & \textcolor{OliveGreen}{green}  \\
Low+Intermediate peaks  & $\kappa_0 \in [-0.06,0.27] $ & 30 & \textcolor{OliveGreen}{$\times$} & \textcolor{OliveGreen}{green}  \\
Intermediate+High peaks  & $\kappa_0 \in [0.1,0.45] $ & 30 & \textcolor{OliveGreen}{$\bullet$} & \textcolor{OliveGreen}{green}  \\
All peaks  & $\kappa_0 \in [-0.06,0.45] $ & 45 & \textcolor{OliveGreen}{$\blacksquare$} & \textcolor{OliveGreen}{green}  \\ \hline
\bottomrule
\end{tabular}
\end{center}
\caption{Catalog of feature types used in this work, along with the chosen number of bands $N_b$ and the plot legends for Figure \ref{curvingnb}.}
\label{featuretable}
\end{table*}

\begin{figure}
\includegraphics[scale=0.3]{Figures/curving_nb.eps}
\caption{Expectation value of the $w$ variance computed with equation (\ref{estimatorcovariance}) as a function of $1/N_r$, for the features outlined in Table \ref{featuretable}. The line width increases with the number of bands $N_b$ used for each feature.}
\label{curvingnb}
\end{figure}

@@

Figure \ref{wvar_nr} show the scaling of the expectation value of $\h{\sigma}_w^2$ with the number of mocks $N_r$ using the power spectrum and compare it with the known result obtained in \citep{DodelsonSchneider13} and expressed in equation (\ref{ourscaling}). Figure \ref{wvar_ns} shows the effect of varying $N_s$ on the $w$ constraint. Finally Figure \ref{curvingnb} shows....

@@

In this section we discuss our main findings. Figure \ref{ps_pdf} shows that, although different choices of $N_s$ do not seem to affect the power spectrum PDF at large scales (top two panels), there are some qualitative differences on small scales (bottom two panels), on which a shear ensemble built with a small $N_s$ does not have the same statistical behavior of ensembles built with large $N_s$. Looking at the black curves, we see that the $N_s=1$ ensembles exibit some stochastic fluctuations in where the PDF peaks. We need as small as $N_s=2$ to recover the right location parameter for the power spectrum on small scales.

>>>>

\begin{abstract}

Constraining cosmology using Weak Gravitational Lensing consists in comparing measured feature vectors, of various dimensions $N_b$, with simulated ones. An accurate estimate of the $N_b\times N_b$ feature covariance matrix is essential to obtain accurate parameter confidence intervals. When this covariance matrix $\bb{C}$ is measured from simulations, an important question to ask is how big the simulation set should be for an accurate estimation of $\bb{C}$. We construct different mock ensembles with $N_r$ realizations of the same shear field, using a common randomization procedure based on $N_s<N_r$ independent $N$--body simulations. Previous work \citep{DodelsonSchneider13} showed that noise fluctuations in the feature covariance lead, at quadratic order, to a $O(1/N_r)$ degradation of parameter constraints. Using a variety of features that include shear-shear power spectra and peak counts, we show that cubic and quartic covariance fluctuations lead to additional $O(1/N_r^2)$ error degradation that is not negligible when $N_r$ is only a factor of a few bigger than $N_b$. When we study the large $N_r$ limit we find that a single $N$--body simulation can be recycled a large number of times to produce an ensemble of $N_r=10^5$ shear realizations that are mutually independent. We also find that a small number of independent $N$--body simulations (1 or 2) are necessary to forecast parameter error bars at percent accuracy, while a larger number might be needed to measure accurate feature ensemble means.

\end{abstract}

@@

\subsection{Shear field simulations}
\label{shearsim}
%
In this paragraph we describe how we constructed our shear field ensembles. Background galaxies at redshift $z_s$ are lensed by large scale structure between $z=0$ and $z_s$. The shape distortions due to the cosmic shear $\pmb{\gamma}$ can be computed in terms of the dark matter gravitational potential $\Phi(\bb{x},z)$. Because the evolution of $\Phi$ in redshift is non--linear, its evolution needs to be computed with numerical simulations. We make use of the public code \texttt{Gadget2} \citep{Gadget2}, with which we run a sequence of dark--matter--only $N$--body simulations that track the evolution of the density fluctuations in the large scale structure of the universe. We assume a standard $\Lambda$CDM framework with $(\Omega_m,\Omega_\Lambda,h,w,\sigma_8,n_s)=(0.26,0.74,0.72,-1,0.8,0.96)$ and we fix the comoving size of the simulation box to $240\mathrm{Mpc}/h$. We fill the box with $512^3$ particles, which correspond to a mass resolution of $10^{10}M_\odot$. We assume a uniform galaxy distribution at a constant redshift $z_s=2$ (at which the simulation box has an angular size of $\theta_{box}=3.5\mathrm{deg}$) and we discretize the mass distribution between $z_s$ and the observer at $z=0$ with a sequence of 46 two dimensional lenses of thickness $80\mathrm{Mpc}/h$. The surface density on each lens is computed by projecting the three dimensional density measured from \texttt{Gadget2} snapshots. We then apply the multi--lens--plane algorithm (see \citep{RayTracingHartlap,RayTracingJain} for example) to trace the deflections of $n_{ray}^2=2048^2$ light rays arranged on a square grid of total size $\theta_{box}$, from $z=0$ to $z_s$. The light deflection calculations have been performed with our implementation of multi--lens--plane algorithm, which is part of the \texttt{LensTools} computing package \citep{LensTools} (released under the \texttt{MIT} license). Different realizations $r$ of the same shear field $\pmb{\gamma}_r(\pmb{\theta})$ can be generated with different choices of the lenses that lay between the observer and $z_s$. The randomization procedure we adopt is the following:

\begin{itemize}
\item Pick a lens redshift $z_l$, and select the snapshot at $z_l$ from the $i$--th $N$--body simulation at disposal, where $i$ is a random number in $[1,N_s]$
\item Choose randomly a direction between the three choices ${\bb{n}_x,\bb{n}_y,\bb{n}_z}$: the lens plane will be pependicular to this direction
\item Choose the cutting point of the plane in the snapshot: because the lens thickness is 1/3 the size of the box, we can cut three different slices of the simulation box for each orientation ${\bb{n}_x,\bb{n}_y,\bb{n}_z}$, which gives a total of 9 choices for generating a lens plane out of a $N$--body snapshot
\item Perform a periodical random shift of the lens plane along its two directions
\item Repeat this procedure for each lens redshift $z_l$  
\end{itemize}  
%
This randomization procedure allows us to produce an arbitrary number $N_r$ of mock shear realizations $\pmb{\gamma}_r(\pmb{\theta})$, which however will not be all independent if $N_s$ is not large enough. Using a set of 200 independent $N$--body simulations, we construct different ensembles with different choices of $N_s\in[1,200]$, each made of 1000 mock shear realizations. We also build an additional ensemble with $N_s=1$ and $10^5$ mocks. From each of these mocks we reconstruct the convergence $\kappa_r(\pmb{\theta})$ and measure its angular power spectrum $P^{\kappa\kappa}_r(l)$ defined as
\begin{equation}
\langle\tilde{\kappa}_r(\bb{l})\tilde{\kappa}_r(\bb{l}')\rangle = (2\pi)^2\delta_D(\bb{l}+\bb{l}')P^{\kappa\kappa}_r(l)
\end{equation}
%
As an additional summary statistic, we consider the counts of local $\kappa$ maxima of a certain height $\kappa_0$, $n_r(\kappa_0)$ (from now on \textit{peak counts}), with varying $\kappa_0$. The fact that the ensemble of $N_r$ mocks is not completely independent if $N_s$ is not large enough can have an effect on the covariance estimators of $P^{\kappa\kappa},n(\kappa_0)$.

@@

\begin{equation}
\label{covest}
\bbh{C} = \frac{1}{N_r-1}\sum_{r=1}^{N_r} (\bbh{d}_r - \bar{\bb{d}}) (\bbh{d}_r - \bar{\bb{d}})^T
\end{equation}
%
Assuming a normal feature likelihood, together with a flat prior on the parameter space, the parameter posterior distribution $\mathcal{L}(\bb{p}\vert\bbh{d}_{obs})$ given an observed instance of $\bbh{d}$, which we call $\bbh{d}_{obs}$, can be obtained using Bayes theorem
\begin{equation}
\label{posteriorbayes}
-2\log\mathcal{L}(\bb{p}\vert\bbh{d}_{obs}) = (\bbh{d}_{obs}-\bb{d}(\bb{p}))^T\bbh{C}^{-1}(\bbh{d}_{obs}-\bb{d}(\bb{p}))
\end{equation}
%
For the sake of simplicity, we approximate the posterior as a Gaussian around its maximum. This is easily done approximating the simulated feature at first order around a point $\bb{p}_0$ (that ideally is the maximum of (\ref{posteriorbayes}))
\begin{equation}
\bb{d}(\bb{p}) \approx \bb{d}_0 + \bb{d}_0^\prime(\bb{p}-\bb{p}_0) 
\end{equation}
%
To measure the feature derivatives $\bb{d}'_0$ with respect to cosmology, we make use of the public code \texttt{NICAEA} \citep{Nicaea} for the power spectrum, and we use an independent simulation set (that contains mocks at different combinations of $(\Omega_m,w,\sigma_8)$, see Table \ref{simtable}) for the peak counts.

\begin{table}
\begin{center}
\begin{tabular}{c|c|c||c|c}
\toprule
$\Omega_m$ &  $w$ & $\sigma_8$ & $N_s$ & Number of $\kappa$ ensembles \\ \hline \hline
\midrule

0.26 & -1 & 0.8 & $1\div 200$ & 16 \\
0.29 & -1 & 0.8 & 1 & 1 \\
0.26 & -0.8 & 0.8 & 1 & 1 \\
0.26 & -1 & 0.6 & 1 & 1 \\

\bottomrule
\end{tabular}
\end{center}
\caption{Summary table of the shear ensembles used in this work.}
\label{simtable}
\end{table}

We can build the estimator for the posterior maximum $\bbh{p}$, given the observation $\bbh{d}_{obs}$, as follows:
%
\begin{equation}
\label{estimatormean}
\bbh{p} = \bb{p}_0 + \bbh{T}(\bbh{d}_{obs}-\bb{d}_0)
\end{equation}

@@

\begin{equation}
\label{estimatorfisher}
\bbh{F} = \bb{d}_0'^T\bbh{\Psi}\bb{d}_0'
\end{equation}
%
Where we introduced the Fisher matrix estimator $\bbh{F} = \bb{d}_0'^T\bbh{\Psi}\bb{d}_0'$ and, for simplicity, we assumed $\langle\bbh{d}_{obs}\rangle_O=\bb{d}_0$, so that $\langle(\bbh{d}_{obs}-\bb{d}_0)(\bbh{d}_{obs}-\bb{d}_0)^T\rangle_O=\bb{C}$. 
When we perform an observation $\bbh{d}_{obs}$, the parameter estimate $\bbh{p}$ is a random draw from a probability distribution with variance $\h{\Sigma}_\bb{p}$, which inherits noise from the simulations. The noise in the covariance estimator (\ref{covest}) and its inverse $\bbh{\Psi}$ propagates all the way to the posterior (\ref{posteriorbayes}), the parameter estimate (\ref{estimatormean}) and its variance (\ref{estimatorcovariance}). Following \citep{DodelsonSchneider13,Taylor12,MasumotoWishart} we compute the expectation value of (\ref{estimatorcovariance}) over simulations, $\langle\h{\Sigma}_\bb{p}\rangle$, up to $O(1/N_r^2)$ by expanding (\ref{estimatorcovariance}) to quartic order in the $\bbh{\Psi}$ statistical fluctuations. Indicating the noiseless parameter covariance as $\Sigma_\bb{p}=\bb{F}^{-1}$, the result we get is \footnote{The details of the calculation are given in Appendix A}

\begin{widetext}
\begin{equation}
\label{quarticdegradation}
\langle\h{\Sigma}_\bb{p}\rangle = \Sigma_\bb{p}\left[1+\frac{N_b-N_p}{N_r}+\frac{(N_b-N_p)(N_b-N_p+2)}{N_r^2}\right] + O\left(\frac{1}{N_r^3}\right)
\end{equation}
\end{widetext}
%
Next, we restrict ourselves to the large $N_r$ limit, and we further investigate the behavior of the $O(1/N_r)$ term. We consider three cases \footnote{The details of these calculations are given in Appendix B}:
\begin{enumerate}
\item If the true data covariance $\bb{C}$ is known, the estimator (\ref{estimatorcovariance}) is biased, and the dominant contribution of the bias comes from the second order fluctuations in $\bbh{\Psi}$. Once the expectation values over simulations are taken, the bias sums up to

\begin{equation}
\label{dodelsonscaling}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}\left(1+\frac{N_b-N_p}{N_r}\right)
\end{equation}
%
This is the result obtained by \citep{DodelsonSchneider13}.

\item Usually the true covariance is not known, and it is tempting to plug in its estimator $\bbh{C}$, measured from the same simulation set we use to compute $\bbh{\Psi}$. This approach has been used before in the literature (see \citep{MinkPetri,MinkShirasaki} for example). If we do that, and we do not correct for the $\bbh{\Psi}$ bias, the parameter variance will have a contribution from both the second and first order fluctuations in $\bbh{\Psi}$, which now have a non zero expectation value. In this case the bias sums up to 
\begin{equation}
\label{mockscalinguncorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}\left(1-\frac{N_b-N_p}{N_r}\right)
\end{equation}

@@

\begin{equation}
\label{mockscalingcorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}\left(1+\frac{1+N_p}{N_r}\right)
\end{equation}

\end{enumerate} 
%
The error degradation  in each parameter $p$, at dominant order, scales as $D/N_r$, where $D=N_b-N_p,N_p-N_b,1+N_p$ for cases 1,2,3 respectively. This means that, when parameter errors are estimated purely from simulations, and the same simulation set is used to estimate both $\bbh{C},\bbh{\Psi}$, these errors tend to decrease with $N_b$ (case 2), or at best remain constant (case 3), while the distribution of $\bbh{p}$ has still a variance that increases with $N_b$. This typically leads to underestimates in parameter errors. We test the scaling relation
\begin{equation}
\label{ourscaling}
\langle\h{\sigma}_p^2\rangle = \sigma^2_{p,\infty}(N_s)\left(1+\frac{D}{N_r}\right)
\end{equation}
%
on our simulations for high and small $N_r$. We use notation $\h{\sigma}^2_p=\h{\Sigma}_{pp}$ and we indicate as $\sigma^2_{p,\infty}$ the expectation value of the variance in the limit of infinite mocks $N_r\rightarrow\infty$. We call $D$ the \textit{effective dimensionality} of the feature space (which, as seen before, can be negative in some cases). We compute the expectation values of $\h{\sigma}^2_p$ (equations (\ref{estimatorcovariance}),(\ref{ourscaling})) averaging over 100 random resamplings of our shear ensembles. For the true feature covariance matrix $\langle(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\rangle=\bb{C}$ we use the estimated covariance from a grand ensemble built with the union of all the ensembles with different $N_s$.

The true parameter variance $\sigma^2_{p,\infty}(N_s)$ in principle depends on the number of independent $N$--body simulations $N_s$, which appears in the randomization procedure we describe in \S~\ref{shearsim}. The reason for this can be found in that, if $N_s$ is not big enough, the different shear realizations cannot be all independent, and hence the true variance $\sigma^2_{p,\infty}(N_s\rightarrow\infty)$ cannot be recovered for low $N_s$ even if $N_r$ is large. In the next paragraph we outline our main findings.

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/means_ns.eps}
\caption{Means of shear features $d$ measured from different shear ensembles. The quantities shown are the differences between mean features measured from ensemble with varying $N_s$, with respect to the means measured from the $N_s=200$ ensemble, in units of the statistical error measured from the $N_s=100$ ensemble. The colored lines refer to shear-shear power spectra measured at $l=115$ (black), 1027 (cyan), 5283 (green) and peak counts of height $\kappa_0=0.05$ (red), 0.17 (purple), 0.28 (orange).}
\label{means_ns}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/curving_nb.eps}
\caption{Expectation value of the $w$ variance computed with equation (\ref{estimatorcovariance}) as a function of $1/N_r$, for the features outlined in Table \ref{featuretable}. The dashed and solid lines show the the theory predictions from equation (\ref{quarticdegradation}) at orders $O(1/N_r),O(1/N_r^2)$. The asymptotic variance $\sigma^2_{w,\infty}$ has been computed from a linear regression of $\langle\h{\sigma}^2_w\rangle$ vs $1/N_r$ for $N_r>500$.}
\label{curvingnb}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_ns.eps}
\caption{Variance on $w$ in the limit $N_r\rightarrow\infty$, varying the mock ensemble used to compute the covariance estimator (\ref{covest}). We show the dependence of $\sigma_{w,\infty}^2(N_s)$ in units of the mean over $N_s$, for the power spectrum logarithmically binned (black), the power spectrum linearly binned (red) and the peak counts (green).}
\label{wvar_ns}
\end{figure}

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}
\toprule
\textbf{Feature} &  \textbf{Specifications} & $N_b$ &  \textbf{Symbol} & \textbf{Color} \\ \hline \hline
\midrule
Power Spectrum, log binning  & $l \in [100,800] $ & 8 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [1000,6000] $ & 7 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [100,6000] $ & 15 & \textcolor{red}{$\times$} & \textcolor{red}{red}  \\
Power Spectrum, linear binning  & $l \in [100,2000] $ & 15 & \textcolor{red}{$+$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $l \in [2500,4500] $ & 15 & \textcolor{red}{$\times$} & \textcolor{red}{red}  \\
Power Spectrum, linear binning  & $l \in [100,4500] $ & 30 & \textcolor{OliveGreen}{$\bullet$} & \textcolor{OliveGreen}{green}  \\ 
Power Spectrum, linear binning  & $l \in [100,6000] $ & 39 & \textcolor{blue}{$\bullet$} & \textcolor{blue}{blue}  \\ \hline
Low peaks  & $\kappa_0 \in [-0.06,0.09] $ & 15 & \textcolor{red}{$+$} & \textcolor{red}{red}  \\ 
Intermediate peaks  & $\kappa_0 \in [0.1,0.27] $ & 15 & \textcolor{red}{$\bigstar$} & \textcolor{red}{red}  \\ 
High peaks  & $\kappa_0 \in [0.28,0.45] $ & 15 & \textcolor{red}{$\diamond$} & \textcolor{red}{red}  \\
Low+Intermediate peaks  & $\kappa_0 \in [-0.06,0.27] $ & 30 & \textcolor{OliveGreen}{$\times$} & \textcolor{OliveGreen}{green}  \\
Intermediate+High peaks  & $\kappa_0 \in [0.1,0.45] $ & 30 & \textcolor{OliveGreen}{$\bullet$} & \textcolor{OliveGreen}{green}  \\
All peaks  & $\kappa_0 \in [-0.06,0.45] $ & 45 & \textcolor{magenta}{$\blacksquare$} & \textcolor{magenta}{magenta}  \\ \hline
\bottomrule
\end{tabular}
\end{center}
\caption{Catalog of feature types used in this work, along with the chosen number of bands $N_b$ and the plot legends for Figure \ref{curvingnb}.}
\label{featuretable}
\end{table*}

@@

Figure \ref{curvingnb} shows a comparison between the dependence of $\langle\h{\sigma^2_w}\rangle$ measured from simulations and the theory prediction from equation (\ref{quarticdegradation}).

Figure \ref{wvar_nr} shows the scaling of the expectation value of $\h{\sigma}_w^2$ when the number of mocks is large, $N_r\gg500$, using the power spectrum and compare it with the known result obtained in \citep{DodelsonSchneider13} and expressed in equation (\ref{ourscaling}). Finally, Figure \ref{wvar_ns} shows the effect of varying $N_s$ on the $w$ constraint.

@@

In this section we discuss our main findings. Figure \ref{ps_pdf} shows that, although different choices of $N_s$ do not seem to affect the power spectrum PDF at large scales (top two panels), there are some qualitative differences on small scales (bottom two panels), on which a shear ensemble built with a small $N_s$ does not have the same statistical behavior of ensembles built with large $N_s$. Looking at the black curves, we see that the $N_s=1$ ensembles exibit some stochastic fluctuations in where the PDF peaks. We need as small as $N_s=2$ to recover the right location parameter for the power spectrum on small scales. Figure \ref{means_ns} shows that mutiple independent $N$--body simulations $N_s>1$ are necessary for measuring the means of feature ensembles to an accuracy corresponding to 10\% of the statistical error. The number of required simulations $N_s$ depends on the feature type and ranges from a few (1 or 2) for the power spectrum at small multipoles $l\lesssim 500$ to an order of $30\div50$ for the power spectrum at large multipoles $l\gtrsim1000$ peak counts at a high threshold $\kappa_0\approx0.3$.
